Historical Soil Organic Carbon Budget

### 2020-01-22 SOCbudgetPaper/SOCbudget.Rmd
several Updates + first part methods

---
title: Historical Soil Organic Carbon Budget
journal: "`r rticles::copernicus_journal_abbreviations(journal_name = 'communication')`"
author:
  - given_name: Kristine
    surname: Karstens
    affiliation: 1
    email: kristine.karstenst@pik-potsdam.de
    corresponding: true
  - given_name: Benjamin Leon
    surname: Bodirsky
    affiliation: 1
  - given_name: Alexander
    surname: Popp
    affiliation: 1

affiliation:
  - code: 1
    address: Potsdam-Institut of Climate Impacts Research, Potsdam, Germany

abstract: |

 SOC one of larges c sinks on earth (3 times larger biospehre pool). Agricultural management leads to a depletion of soil organic crabon. However
 this depletion of soil organic carbon (SOC) pools are so far not well represented in global assessments of historic carbon emissions. While SOC
 models often represent well the biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these 
 biophysical processes are still little investigated.
 Here we create a spatial explicit data set for crop residue and manure management on cropland based on global historic production (FAOSTAT) and 
 land-use (LUH2) data and combine it with the IPCC Tier 2 approach to create a half-degree resolution soil organic carbon budget on mineral soils.
 We estimate that due to arable farming soils have lost over (?) GtOC of which (??) GtOC have been released within the period 1990-2010.
 Tier 2 IPCC methodolgy estimates higher soil organic carbon losses than Tier 1 methods, which may origin from … . We also find that SOC is very 
 sensity to management decision such as residue recycling indicating the nessessity to incorporated better management data in soil models.

bibliography: SOCbudget.bib
running:
  title: R Markdown Template for Copernicus
  author: Nüst et al.
# This section is mandatory even if you declare that no competing interests are present.
competinginterests: |
  The authors declare no competing interests.
# OPTIONAL:

algorithms: true
# See https://publications.copernicus.org/for_authors/licence_and_copyright.html, normally used for transferring the copyright, if needed.

copyrightstatement: |
  The author's copyright for this publication is transferred to institution/company.

availability:
  codedata: |
    use this to add a statement when having data sets and software code available

authorcontribution: |
  Karstens wrote code and paper build on work of Bodirsky. Bodirsky and Popp revised paper.

disclaimer: |
  We like Copernicus.

acknowledgements: |
  Thanks to the rticles contributors!

appendix: |
  \section{Figures and tables in appendices}
  Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:
  \subsection{Option 1}
  If you sorted all figures and tables into the sections of the text, please also sort the appendix figures and appendix tables into the respective appendix sections.
  They will be correctly named automatically.
  \subsection{Option 2}
  If you put all figures after the reference list, please insert appendix tables and figures after the normal tables and figures.

  To rename them correctly to A1, A2, etc., please add the following commands in front of them:
  `\appendixfigures` needs to be added in front of appendix figures
  `\appendixtables` needs to be added in front of appendix tables

  Please add `\clearpage` between each table and/or figure. Further guidelines on figures and tables can be found below.

output:
  rticles::copernicus_article: default
  bookdown::pdf_book:
    base_format: rticles::copernicus_article # for using bookdown features like \@ref()
---

>>>>

---
title: Historical Soil Organic Carbon Budget
journal: "`r rticles::copernicus_journal_abbreviations(journal_name = 'communication')`"
author:
  - given_name: Kristine
    surname: Karstens
    affiliation: 1
    email: kristine.karstenst@pik-potsdam.de
    corresponding: true
  - given_name: Benjamin Leon
    surname: Bodirsky
    affiliation: 1
  - given_name: Alexander
    surname: Popp
    affiliation: 1
affiliation:
  - code: 1
    address: Potsdam-Institut of Climate Impacts Research, Potsdam, Germany
abstract: | 
 SOC one of larges c sinks on earth (3 times larger biospehre pool). Agricultural management leads to a depletion of soil organic crabon. However
 this depletion of soil organic carbon (SOC) pools are so far not well represented in global assessments of historic carbon emissions. While SOC
 models often represent well the biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these 
 biophysical processes are still little investigated.
 Here we create a spatial explicit data set for crop residue and manure management on cropland based on global historic production (FAOSTAT) and 
 land-use (LUH2) data and combine it with the IPCC Tier 2 approach to create a half-degree resolution soil organic carbon budget on mineral soils.
 We estimate that due to arable farming soils have lost over (?) GtOC of which (??) GtOC have been released within the period 1990-2010.
 Tier 2 IPCC methodolgy estimates higher soil organic carbon losses than Tier 1 methods, which may origin from … . We also find that SOC is very 
 sensity to management decision such as residue recycling indicating the nessessity to incorporated better management data in soil models. 
bibliography: SOCbudget.bib
running:
  title: R Markdown Template for Copernicus
  author: Nüst et al.
competinginterests: |
  The authors declare no competing interests.
# OPTIONAL:
algorithms: true
# See https://publications.copernicus.org/for_authors/licence_and_copyright.html, normally used for transferring the copyright, if needed.
copyrightstatement: |
  The author's copyright for this publication is transferred to institution/company.
availability:
  codedata: |
    use this to add a statement when having data sets and software code available
authorcontribution: |
  Karstens wrote code and paper build on work of Bodirsky. Bodirsky and Popp revised paper.
disclaimer: |
  We like Copernicus.
acknowledgements: |
  Thanks to the rticles contributors!
appendix: |
  \section{Figures and tables in appendices}
  \subsection{Option 1}
  If you sorted all figures and tables into the sections of the text, please also sort the appendix figures and appendix tables into the respective appendix sections.
  They will be correctly named automatically.
  \subsection{Option 2}
  If you put all figures after the reference list, please insert appendix tables and figures after the normal tables and figures.

  `\appendixfigures` needs to be added in front of appendix figures
  `\appendixtables` needs to be added in front of appendix tables

  Please add `\clearpage` between each table and/or figure. Further guidelines on figures and tables can be found below.
  Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:
  To rename them correctly to A1, A2, etc., please add the following commands in front of them:
output:
  bookdown::pdf_book:
    base_format: rticles::copernicus_article # for using bookdown features like \@ref()
  rticles::copernicus_article: default
  md_document:
    variant: markdown_github
---

### 2020-01-22 SOCbudgetPaper/SOCbudget.tex
several Updates + first part methods

The following settings can or must be configured in the header of this
file and are bespoke for Copernicus manuscripts:

\begin{itemize}
\item
  The \texttt{journal} you are submitting to using the official
  abbreviation.
\item
  You can use the function
  \texttt{rticles::copernicus\_journal\_abbreviations(name\ =\ \textquotesingle{}...\textquotesingle{})}
  to search the existing journals.
\end{itemize}

@@

Based on the IPCC guidelines \citep[\citet{IPCC2019}]{IPCC2006} for soil
organic carbon stocks, we combining approaches to estimate SOC stocks by
weighting inflows via dead plant material (see @ref(sec:carboninputs))
against outflows through SOC decay (see @ref(sec:tier1),
@ref(sec:tier2)). Carbon displacement via leaching and erosion is
neglected for this budget.

Connecting carbon inputs with soil carbon dynamics will create a budget
of in- and outflows.

@@

Yearly turnover between three different SOC pools for the topsoil is
model using global parameters.

@@

Stock changes are calculated using simple change factors on the SOC
pools for the topsoil.

Carbon inputs are estimated based on the land use. Whereas cropland
inputs are mainly formed by residue, dead below ground and cover crop
biomass alterate by the corresponding management practice, pasture and
natural vegetation inputs are estimate via modelled annual litterfall
rates. For the former we use various data sets and rule based
approaches, which are specified in @ref(sec:agrimanagement)

@@

\section{Content section with citations}

See the
\href{http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html}{R
Markdown docs for bibliographies and citations}.

Copernicus supports biblatex and a sample bibliography is in file
\texttt{sample.bib}. Read {[}{]}, and {[}see {]}.

\section{Content section with R code chunks}

You should always use \texttt{echo\ =\ FALSE} on R Markdown code blocks
as they add formatting and styling not desired by Copernicus. The hidden
workflow results in 42.

You can add verbatim code snippets without extra styles by using
\texttt{\textasciigrave{}\textasciigrave{}\textasciigrave{}} without
additional instructions.

\begin{verbatim}
sum <- 1 + 41
\end{verbatim}

\section{Content section with list}

If you want to insert a list, you must

\begin{itemize}
\item
  leave
\item
  empty lines
\item
  between each list item
\end{itemize}

\section{Examples from the official template}

\subsection{FIGURES}

When figures and tables are placed at the end of the MS (article in
one-column style), please add \clearpage between bibliography and first
table and/or figure as well as between each table and/or figure.

\subsection{FIGURES}

When figures and tables are placed at the end of the MS (article in
one-column style), please add \clearpage between bibliography and first
table and/or figure as well as between each table and/or figure.

\subsubsection{ONE-COLUMN FIGURES}

Include a 12cm width figure of Nikolaus Copernicus from
\href{https://en.wikipedia.org/wiki/File:Nikolaus_Kopernikus.jpg}{Wikipedia}
with caption using R Markdown.

\subsubsection{TWO-COLUMN FIGURES}

You can also include a larger figure.

\subsection{TABLES}

You can ad \LaTeX table in an R Markdown document to meet the template
requirements.

\subsubsection{ONE-COLUMN TABLE}

\begin{table}[t]
\caption{TEXT}
\begin{tabular}{l c r}
\tophline

a & b & c \\
\middlehline
1 & 2 & 3 \\

\bottomhline
\end{tabular}
\belowtable{Table Footnotes}
\end{table}

\subsubsection{TWO-COLUMN TABLE}

\begin{table*}[t]
\caption{TEXT}
\begin{tabular}{l c r}
\tophline

a & b & c \\
\middlehline
1 & 2 & 3 \\

\bottomhline
\end{tabular}
\belowtable{Table footnotes}
\end{table*}

\subsection{MATHEMATICAL EXPRESSIONS}

All papers typeset by Copernicus Publications follow the math
typesetting regulations given by the IUPAC Green Book (IUPAC:
Quantities, Units and Symbols in Physical Chemistry, 2nd Edn., Blackwell
Science, available at:
http://old.iupac.org/publications/books/gbook/green\_book\_2ed.pdf,
1993).

Physical quantities/variables are typeset in italic font (t for time, T
for Temperature)

Indices which are not defined are typeset in italic font (x, y, z, a, b,
c)

Items/objects which are defined are typeset in roman font (Car A, Car B)

Descriptions/specifications which are defined by itself are typeset in
roman font (abs, rel, ref, tot, net, ice)

Abbreviations from 2 letters are typeset in roman font (RH, LAI)

Vectors are identified in bold italic font using \vec{x}

Matrices are identified in bold roman font

Multiplication signs are typeset using the LaTeX commands
\texttt{\textbackslash{}times} (for vector products, grids, and
exponential notations) or \texttt{\textbackslash{}cdot}

The character * should not be applied as mutliplication sign

\subsection{ALGORITHM}

If you want to use algorithms, you can either enable the required
packages in the header (the default, see \texttt{algorithms:\ true}), or
make sure yourself that the \LaTeX packages \texttt{algorithms} and
\texttt{algorithmicx} are installed so that \texttt{algorithm.sty}
respectively \texttt{algorithmic.sty} can be loaded by the Copernicus
template. Copernicus staff will remove all undesirable packages from
your LaTeX source code, so please stick to using the header option,
which only adds the two acceptable packages.

\begin{algorithm}
\caption{Algorithm Caption}
\label{a1}
\begin{algorithmic}
\STATE $i\gets 10$
  \IF {$i\geq 5$} 
\STATE $i\gets i-1$
  \ELSE
\IF {$i\leq 3$}
\STATE $i\gets i+2$
  \ENDIF
\ENDIF
\end{algorithmic}
\end{algorithm}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
\section{Figures and tables in appendices}

Regarding figures and tables in appendices, the following two options
are possible depending on your general handling of figures and tables in
the manuscript environment: \subsection{Option 1} If you sorted all
figures and tables into the sections of the text, please also sort the
appendix figures and appendix tables into the respective appendix
sections. They will be correctly named automatically.
\subsection{Option 2} If you put all figures after the reference list,
please insert appendix tables and figures after the normal tables and
figures.

To rename them correctly to A1, A2, etc., please add the following
commands in front of them: \texttt{\textbackslash{}appendixfigures}
needs to be added in front of appendix figures
\texttt{\textbackslash{}appendixtables} needs to be added in front of
appendix tables

Please add \texttt{\textbackslash{}clearpage} between each table and/or
figure. Further guidelines on figures and tables can be found below.
\noappendix

>>>>

@@

Based on the IPCC guidelines vol.~4 and their refinement
(\citep{ipcc_2006_2006},\citep{ipcc_2019_2019}) for soil organic carbon
stocks, we combining approaches to estimate SOC stocks by weighting
inflows via dead plant material (see \ref{sec:carboninputs}) against
outflows through SOC decay (see \ref{sec:tier2}). Carbon displacement
via leaching and erosion is neglected in this study. We calculate annual
land use type specific budgets for cropland, pastures and natural
vegetation, also representin land conversion as transfer between landuse
type budgets. A simple approach based o the tier 1 method of the older
IPCC guidelines vol 4. (\citet{ipcc_2006_2006}) using stock change
factors, is applied to cross validated results \ref{sec:tier1}

Carbon input estimations are based on the land use type. Whereas
cropland inputs are mainly formed by disaggregated country statistics on
residue, dead below ground and cover crop biomass, pasture and natural
vegetation inputs are estimate via modelled annual litterfall rates.
Using the steady-state method of the IPCC guidelines
(\citep{ipcc_2019_2019}) carbon inputs have to be accompined by data on
lignin and nitrogen content to allocate dead plant biomass to the
corrosponding soil pools based on the chemical texture. Sources for all
use data can be found in table \ref{tab:datasourceinputs}

\begin{table*}[h]
\caption{Sources for carbon input data}
\begin{tabular}{l l l}
\tophline
 Land use types   & carbon inputs & nitrogen and lignin content \\
\middlehline
 Cropland         & FAO statistics, AQUASTAT, LPJmL4 [1] & default values [\cite{ipcc_2006_2006}] \\
 Pasture          & annual litterfall in $\tfrac{gC}{m^2}$ from LPJmL4 - manage grassland [3] & default values [2] \\
 Natural vegetation & annual litterfall in $\tfrac{gC}{m^2}$ from LPJmL4  & \begin{minipage}[t]{0.37\columnwidth}\raggedright\strut Nitrogen and lignin content of tree compartments used in CENTURY \strut \end{minipage}\tabularnewline
\bottomhline
\end{tabular}
\label{tab:datasourceinputs}
\belowtable{}
\end{table*}

@@

We are following the steady-state method of the refinement of the IPCC
guidelines vol.~4 (\citet{ipcc_2019_2019}) by calculating yearly
turnover and transfer rates between three different SOC pools for the
topsoil (0-30 cm). The approach is based on global parameters
({[}@ref(ipcc\_2019\_2019){]}) as well as half-degree data on sand
fraction (SoilGrids), temperature, preciptation and potential
evapotranspiration (CRU). Next to the given climatic and natural
biophysical properties irrigation regime (rainfed vs.~irrigated) as well
as tillage (as soil disturbance indicator) modificate processes. For
cropland an assessment of tillage types and irrigation conditions has
been made, whereas on pastures and natural vegetation, we assume rainfed
and non-tilled conditions.

@@

Following the tier 1 approach of the IPCC guidelines vol.~4
(\citet{ipcc_2006_2006}), stocks are estimated via stock change factors
given by the IPCC for the topsoil (0-30 cm) and based on measurements.
The factors are differentiate between different crop and management
systems reflecting different dynamics under changed in- and outflows
without explicitly tracking these. They can be seen as conservative
guesses and will be used to evaluate our modelling based results.

@@

\section{Results}

\section{Discussion}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
\section{Figures and tables in appendices}\subsection{Option 1}

If you sorted all figures and tables into the sections of the text,
please also sort the appendix figures and appendix tables into the
respective appendix sections. They will be correctly named
automatically. \subsection{Option 2} If you put all figures after the
reference list, please insert appendix tables and figures after the
normal tables and figures.

\texttt{\textbackslash{}appendixfigures} needs to be added in front of
appendix figures \texttt{\textbackslash{}appendixtables} needs to be
added in front of appendix tables

Please add \texttt{\textbackslash{}clearpage} between each table and/or
figure. Further guidelines on figures and tables can be found below.
Regarding figures and tables in appendices, the following two options
are possible depending on your general handling of figures and tables in
the manuscript environment: To rename them correctly to A1, A2, etc.,
please add the following commands in front of them:
\noappendix
